---
title: "wx_station_data"
author: "Erika Lee"
date: "2023-11-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#installing packages / set working directory
packages <- c('tidyverse',
              'lterdatasampler', 'dplyr', 'ggplot2', 'readr', 'readxl', 'lubridate', 'plotly', 'openxlsx', 'plotly', 'rstatix')

#install.packages(packages)

library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

## Tree Temp Data

### 105E Pers_b Tree Temp Data

```{r}
#read in temperature data
pers_b_tree_composite=read_excel("nsf/105E_pers_burned/105E_pers_b_composite.xlsx")

#filter temp data to full data frame
pers_b_tree_fulldata <- pers_b_tree_composite %>%
  mutate(datetime = ymd_hms(TIMESTAMP)) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, RECORD))

#adding in an aspect column - WITH ONLY 14 SENSORS!
pers_b_aspects <- pers_b_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(5)",
      "Temp_C_Avg(9)",
      "Temp_C_Avg(13)", "Temp_C_Avg(15)"
    ),
    "north",
    "other"
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(2)",
      "Temp_C_Avg(6)",
      "Temp_C_Avg(10)",
      "Temp_C_Avg(14)", "Temp_C_Avg(16)"),
    "south",
    aspect
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(3)", "Temp_C_Avg(7)", "Temp_C_Avg(11)"),
    "charred",
    aspect
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(4)", "Temp_C_Avg(8)", "Temp_C_Avg(12)"),
    "uncharred",
    aspect
  )) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(2)",
      "Temp_C_Avg(3)",
      "Temp_C_Avg(4)"
    ),
    "gb_w",
    "other"
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(5)",
      "Temp_C_Avg(6)",
      "Temp_C_Avg(7)",
      "Temp_C_Avg(8)"
    ),
    "gb_e",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(9)",
      "Temp_C_Avg(10)",
      "Temp_C_Avg(11)",
      "Temp_C_Avg(12)"
    ),
    "gb_s",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(13)", "Temp_C_Avg(14)"),
    "db_n",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(15)", "Temp_C_Avg(16)"),
    "db_s",
    tree_name
  ))

#creating dataframes for individual tree
pers_gb_w <- pers_b_aspects %>%
  filter(sensor_id == c("Temp_C_Avg(1)", "Temp_C_Avg(2)", "Temp_C_Avg(3)", "Temp_C_Avg(4)"))

pers_gb_e <- pers_b_aspects %>%
  filter(sensor_id == c("Temp_C_Avg(5)", "Temp_C_Avg(6)", "Temp_C_Avg(7)", "Temp_C_Avg(8)"))

pers_gb_s <- pers_b_aspects %>%
  filter(sensor_id == c("Temp_C_Avg(9)", "Temp_C_Avg(10)", "Temp_C_Avg(11)", "Temp_C_Avg(12)"))

pers_db_n <- pers_b_aspects %>%
  filter(sensor_id == c("Temp_C_Avg(13)", "Temp_C_Avg(14)"))

pers_db_s <- pers_b_aspects %>%
  filter(sensor_id == c("Temp_C_Avg(15)", "Temp_C_Avg(16)"))

#exporting aspect data as excel

write.xlsx(pers_b_aspects, "nsf/105E_pers_burned/105E_pers_b_aspects.xlsx")
```

### 105E Pers_b Plots

```{r}
#colors for sensors
color_data_pers <- data.frame(tree_name = c("gb_e", "gb_w", "gb_s", "db_n", "db_s"), color = c("blue", "orange", "red", "black", "purple"))

#individual tree timeseries plots
pers_north <- pers_b_aspects %>%
  filter(aspect == "north") %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = tree_name, alpha = 0.5)) +
  labs(x = "Datetime", y = "Temp (C)", title = "Pers North Sensors") +
  scale_color_manual(values = setNames(color_data_pers$color, color_data_pers$tree_name))

ggplotly(pers_north)

pers_south <- pers_b_aspects %>%
  filter(aspect == "south") %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = tree_name, alpha = 0.5)) +
  labs(x = "Datetime", y = "Temp (C)", title = "Pers South Sensors") +
  scale_color_manual(values = setNames(color_data_pers$color, color_data_pers$tree_name))

ggplotly(pers_south)

pers_charred <- pers_b_aspects %>%
  filter(aspect == "charred") %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = tree_name, alpha = 0.5)) +
  labs(x = "Datetime", y = "Temp (C)", title = "Pers Charred Sensors") +
  scale_color_manual(values = setNames(color_data_pers$color, color_data_pers$tree_name))

ggplotly(pers_charred)

pers_uncharred <- pers_b_aspects %>%
  filter(aspect == "uncharred") %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = tree_name, alpha = 0.5)) +
  labs(x = "Datetime", y = "Temp (C)", title = "Pers Uncharred Sensors") +
  scale_color_manual(values = setNames(color_data_pers$color, color_data_pers$tree_name))

ggplotly(pers_uncharred)

##pers_gb_w plot
pers_gb_w_timeseries <- pers_gb_w %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  #facet_wrap(~aspect) + 
  labs(x = "Datetime", y = "Temp (C)", title = "pers_gb_w")
  #scale_x_datetime(limits = c(as.POSIXct("2023-11-03 00:00:00"), as.POSIXct("2023-11-10 00:00:00")))

ggplotly(pers_gb_w_timeseries)

##pers_gb_e plot
pers_gb_e_timeseries <- pers_gb_e %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  #facet_wrap(~aspect) + 
  labs(x = "Datetime", y = "Temp (C)", title = "pers_gb_e")
  #scale_x_datetime(limits = c(as.POSIXct("2023-11-03 00:00:00"), as.POSIXct("2023-11-10 00:00:00")))
 
ggplotly(pers_gb_e_timeseries)

##pers_gb_s plot
pers_gb_s_timeseries <- pers_gb_s %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  #facet_wrap(~aspect) + 
  labs(x = "Datetime", y = "Temp (C)", title = "pers_gb_s")
  #scale_x_datetime(limits = c(as.POSIXct("2023-11-03 00:00:00"), as.POSIXct("2023-11-10 00:00:00")))
 
ggplotly(pers_gb_s_timeseries)

##pers_db_n plot
pers_db_n_timeseries <- pers_db_n %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  #facet_wrap(~aspect) + 
  labs(x = "Datetime", y = "Temp (C)", title = "pers_db_n")
  #scale_x_datetime(limits = c(as.POSIXct("2023-11-03 00:00:00"), as.POSIXct("2023-11-10 00:00:00")))
 
ggplotly(pers_db_n_timeseries)

#pers_db_s plot
pers_db_s_timeseries <- pers_db_s %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  #facet_wrap(~aspect) + 
  labs(x = "Datetime", y = "Temp (C)", title = "pers_db_s")
  #scale_x_datetime(limits = c(as.POSIXct("2023-12-06 14:00:00"), as.POSIXct("2023-12-21 13:15:00")))

ggplotly(pers_db_s_timeseries)

#timeseries
pers_timeseries <- pers_b_aspects %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  facet_wrap(~sensor_name) + 
  labs(x = "Datetime", y = "Temp (C)", title = "pers_b_timeseries") +
  #datetime scale, so you can see patterns better
  scale_x_datetime(limits = c(as.POSIXct("2023-11-03 00:00:00"), as.POSIXct("2023-11-10 00:00:00")))

print(pers_timeseries)

#timeseries plot
pers_b_timeseries <- pers_b_aspects %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp)) + facet_wrap("aspect")

plot(pers_b_timeseries)

#histogram - all sensors
pers_b_histogram <- pers_b_aspects %>%
  ggplot(aes(x = temp)) +
  geom_histogram() + facet_wrap( ~ aspect, scale = "free_x") +
  labs(title = "Histogram of Temperature Across Sites",
       x = "Temperature (C)",
       y = "Frequency") +
  theme_minimal()
    # geom_vline(aes(xintercept = mean_temp), price_stats, color = "red", linewidth = 2)

print(pers_b_histogram)
```

### 105E Pers_b Data Exploration

```{r}
#pers_north_GB tree temp data
#view the dataset - must use data() function to bring it into the environment
pers_n_gb_temp <- pers_b_aspects %>%
  filter(aspect == "north") %>%
  filter(tree_name %in% c("gb_w", "gb_s", "gb_e")) %>%
  drop_na()
  #filter(aspect == "north" & datetime >= "2023-12-07 00:00:00" & datetime <= "2024-01-10 00:00:00")

#glimpse is a different way to view dataset, a bit more clear
glimpse(pers_n_gb_temp)

#create a new variable to test 
pers_n_gb_clean<- pers_n_gb_temp %>%
  filter(aspect == "north")

#visualize the relationship
pers_n_gb_clean %>%
  ggplot(aes(x = tree_name, y = temp)) +
  geom_boxplot()

print(pers_n_gb_clean)

#check p-value and assumptions
pers_n_gb_clean %>%
  levene_test(temp ~ tree_name)
#this gives me a p-value of 0.643, which is not significant, which means there is not a difference in variances noted... 
```

```{r}
#create a new variable to test
pers_n_gb_clean<- pers_n_gb_temp %>%
  filter(aspect == "north")

#visualize the relationship
pers_n_gb_clean %>%
  ggplot(aes(x = tree_name, y = temp)) +
  geom_boxplot()

print(pers_n_gb_clean)

#check p-value and assumptions
pers_n_gb_clean %>%
  levene_test(temp ~ tree_name)
#this gives me a p-value of 0.643, which is not significant, which means there is not a difference in variances noted... 
```

```{r}
#pers_n_db tree temp comparison
pers_n_db_temp <- pers_b_aspects %>%
  filter(aspect == "north",tree_name %in% c("db_n", "db_s")) %>%
  drop_na()
  #filter(aspect == "north" & datetime >= "2023-12-07 00:00:00" & datetime <= "2024-01-10 00:00:00")

glimpse(pers_n_db_temp)

pers_n_db_clean<- pers_n_db_temp

pers_n_db_clean %>%
  ggplot(aes(x = tree_name, y = temp)) +
  geom_boxplot()

print(pers_n_db_clean)

#check p-value and assumptions
pers_n_db_clean %>%
  levene_test(temp ~ tree_name)

#not a significant p-value, 0.771
```

```{r}
#pers_s_gb tree temp data
pers_s_gb_temp <- pers_b_aspects %>%
  filter(aspect == "south") %>%
  filter(tree_name %in% c("gb_w", "gb_s", "gb_e")) %>%
  drop_na()
  #filter(aspect == "south" & datetime >= "2023-12-07 00:00:00" & datetime <= "2024-01-10 00:00:00")

#glimpse is a different way to view dataset, a bit more clear
glimpse(pers_s_gb_temp)

#create a new variable to test 
pers_s_gb_clean<- pers_s_gb_temp

#visualize the relationship
pers_s_gb_clean %>%
  ggplot(aes(x = tree_name, y = temp)) +
  geom_boxplot()

print(pers_s_gb_clean)

#check p-value and assumptions
pers_s_gb_clean %>%
  levene_test(temp ~ tree_name)
#gives a p-value of 0.00937
```

Since p-value is statistically significant, variances are not equal. So we will transform the variable with a log.

```{r}
hist(pers_s_gb_clean$temp)

#this part of test did not work...
pers_s_gb_clean %>%
  levene_test(log(temp) ~ tree_name)
```

```{r}
any_missing <- any(is.na(pers_s_temp_clean))

print(any_missing)
#returns false


my_vector <- unlist(pers_s_temp_clean)
any_infinite <- any(is.infinite(my_vector))
any_nan <- any(is.nan(my_vector))
print(any_infinite)  # Check if there are any infinite values
print(any_nan)
```

```{r}
#performing t-test on non-transformed variable

welch_t_test_result <- pers_s_gb_clean %>%
  t_test(temp ~ tree_name, var.equal = FALSE)

# Print the result
print(welch_t_test_result)
#p-value is significant for all tree comparisons
```

```{r}
#pers_s_db tree temp comparison
pers_s_db_temp <- pers_b_aspects %>%
  filter(aspect == "south" & tree_name %in% c("db_n", "db_s")) %>%
  drop_na()
  #filter(aspect == "south" & datetime >= "2023-12-07 00:00:00" & datetime <= "2024-01-10 00:00:00")

glimpse(pers_s_db_temp)

pers_s_db_clean<- pers_s_db_temp

pers_s_db_clean %>%
  ggplot(aes(x = tree_name, y = temp)) +
  geom_boxplot()

print(pers_s_db_clean)

#check p-value and assumptions
pers_s_db_clean %>%
  levene_test(temp ~ tree_name)
#p-value is o.0514 - significant enough... 
```

```{r}
hist(pers_s_db_clean$temp)

#this part of test did not work...
pers_s_db_clean %>%
  levene_test(log(temp) ~ tree_name)
#did not work... so do welsh test...
```

```{r}
#performing t-test on non-transformed variable

welch_t_test_result <- pers_s_db_clean %>%
  t_test(temp ~ tree_name, var.equal = FALSE)

# Print the result
print(welch_t_test_result)
#p-value is not quite significant here... 0.0444
```

Plotting GB_N vs GB_S

```{r}

#creating a new dataframe with name changes to sensors
pers_n_gb_new <- pers_n_gb_temp %>%
  mutate(tree_name = case_when(
    tree_name == "gb_w" ~ "n_gb_w",
    tree_name == "gb_s" ~ "n_gb_s",
    tree_name == "gb_e" ~ "n_gb_e",
    TRUE ~ tree_name  # Keep the original value if it doesn't match the conditions
  ))

pers_s_gb_new <- pers_s_gb_temp %>%
  mutate(tree_name = case_when(
    tree_name == "gb_w" ~ "s_gb_w",
    tree_name == "gb_s" ~ "s_gb_s",
    tree_name == "gb_e" ~ "s_gb_e",
    TRUE ~ tree_name  # Keep the original value if it doesn't match the conditions
  ))

#facet wrapped plots by tree
pers_n_s_gb_plot <- ggplot() +
  geom_line(data = pers_n_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "North"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = pers_s_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "South"), alpha = 0.3) +
  scale_color_manual(values = c("black", "red", "blue", "black", "red", "blue"), name = NULL) +
  scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Pers GB North and South Sensors")
#facet_wrap(~tree_name)

ggplotly(pers_n_s_gb_plot)

```

Plotting DB_N vs DB_S

```{r}
#creating a new dataframe with name changes to sensors
pers_n_db_new <- pers_n_db_temp %>%
  mutate(tree_name = case_when(
    tree_name == "db_n" ~ "n_db_n",
    tree_name == "db_s" ~ "n_db_s",
    TRUE ~ tree_name  # Keep the original value if it doesn't match the conditions
  ))

pers_s_db_new <- pers_s_db_temp %>%
  mutate(tree_name = case_when(
    tree_name == "db_n" ~ "s_db_n",
    tree_name == "db_s" ~ "s_db_s",
    TRUE ~ tree_name  # Keep the original value if it doesn't match the conditions
  ))

#plotting
pers_n_s_db_plot <- ggplot() +
  geom_line(data = pers_n_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "North"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = pers_s_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "South"), alpha = 0.3) +
scale_color_manual(values = c("blue", "red", "blue", "red"), name = NULL) +
  scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Pers DB North and South Sensors")
#facet_wrap(~tree_name)

ggplotly(pers_n_s_db_plot)
```

Plotting Pers_N GB vs DB

```{r}
pers_n_gb_vs_db_plot <- ggplot() +
  geom_line(data = pers_n_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "North"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = pers_n_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "South"), alpha = 0.3) +
  scale_color_manual(values = c("black", "red", "blue", "black", "red", "blue"), name = NULL) +
  #scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Pers North GB vs DB")

ggplotly(pers_n_gb_vs_db_plot)
```

Plotting Pers_S GB vs DB

```{r}
pers_s_gb_vs_db_plot <- ggplot() +
  geom_line(data = pers_s_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "North"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = pers_s_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "South"), alpha = 0.3) +
  scale_color_manual(values = c("black", "red", "blue", "black", "red", "blue"), name = NULL) +
  #scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Pers South GB vs DB")

ggplotly(pers_s_gb_vs_db_plot)
```

```{r}
```

### 105E Trans Burned Tree Temp Data

```{r}
#read in temperature data
trans_b_tree_composite=read_excel("nsf/105E_trans_burned/105E_trans_b_composite.xlsx")

#filter temp data to full data frame
trans_b_tree_fulldata <- trans_b_tree_composite %>%
  mutate(datetime = ymd_hms(TIMESTAMP)) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, Record)) %>%
  na.omit()

#adding in an aspect column - ** need to edit date on trans_b sensors, seems weird, should be 2023.12.06 around 10 am?
trans_b_aspects <- trans_b_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(1)", "Temp_C_Avg(5)", "Temp_C_Avg(9)", "Temp_C_Avg(11)"),
    "north",
    "other"
  )) %>%
mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(2)", "Temp_C_Avg(6)", "Temp_C_Avg(10)", "Temp_C_Avg(12)"),
    "south",
    aspect)) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(3)", "Temp_C_Avg(7)"),
    "charred",
    aspect)) %>%
    mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(4)", "Temp_C_Avg(8)"),
    "uncharred",
    aspect)) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(1)", "Temp_C_Avg(2)", "Temp_C_Avg(3)", "Temp_C_Avg(4)"),
    "gb_n",
    "other"
  )) %>%
mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(5)", "Temp_C_Avg(6)", "Temp_C_Avg(7)", "Temp_C_Avg(8)"),
    "gb_s",
    tree_name)) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(9)", "Temp_C_Avg(10)"),
    "db_n",
    tree_name)) %>%
    mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(11)", "Temp_C_Avg(12)"),
    "db_s",
    tree_name))

#exporting aspect data as excel

write.xlsx(trans_b_aspects, "nsf/105E_trans_burned/105E_trans_b_aspects.xlsx")

#creating dataframes for individual tree
trans_gb_n <- trans_b_aspects %>%
  filter(sensor_id == c("Temp_C_Avg(1)", "Temp_C_Avg(2)", "Temp_C_Avg(3)", "Temp_C_Avg(4)"))

trans_gb_s <- trans_b_aspects %>%
  filter(sensor_id == c("Temp_C_Avg(5)", "Temp_C_Avg(6)", "Temp_C_Avg(7)", "Temp_C_Avg(8)"))

trans_db_n <- trans_b_aspects %>%
  filter(sensor_id == c("Temp_C_Avg(9)", "Temp_C_Avg(10)"))

trans_db_s <- trans_b_aspects %>%
  filter(sensor_id == c("Temp_C_Avg(9)", "Temp_C_Avg(10)"))
```

### 105E Trans_b Plots

```{r}
#colors for sensors
color_data_trans <- data.frame(tree_name = c("gb_n", "gb_s", "db_n", "db_s"), color = c("blue", "orange", "black", "red"))

#individual tree timeseries plots
trans_north <- trans_b_aspects %>%
  filter(aspect == "north") %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = tree_name, alpha = 0.5)) +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans North Sensors") +
  scale_color_manual(values = setNames(color_data_trans$color, color_data_trans$tree_name))

ggplotly(trans_north)

trans_south <- trans_b_aspects %>%
  filter(aspect == "south") %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = tree_name, alpha = 0.5)) +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans South Sensors") +
  scale_color_manual(values = setNames(color_data_trans$color, color_data_trans$tree_name))

ggplotly(trans_south)

trans_charred <- trans_b_aspects %>%
  filter(aspect == "charred") %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = tree_name, alpha = 0.5)) +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans Charred Sensors") +
  scale_color_manual(values = setNames(color_data_trans$color, color_data_trans$tree_name))

ggplotly(trans_charred)

trans_uncharred <- trans_b_aspects %>%
  filter(aspect == "uncharred") %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = tree_name, alpha = 0.5)) +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans Uncharred Sensors") +
  scale_color_manual(values = setNames(color_data_trans$color, color_data_trans$tree_name))

ggplotly(trans_uncharred)

##trans_gb_n plot
trans_gb_n_timeseries <- trans_gb_n %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  facet_wrap(~aspect) + 
  labs(x = "Datetime", y = "Temp (C)", title = "trans_gb_n")
  #scale_x_datetime(limits = c(as.POSIXct("2041-07-16 23:15:00
#T"), as.POSIXct("2041-07-17 02:15:00")))

ggplotly(trans_gb_n_timeseries)

##trans_gb_s plot
trans_gb_s_timeseries <- trans_gb_s %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  facet_wrap(~aspect) + 
  labs(x = "Datetime", y = "Temp (C)", title = "trans_gb_s")
  #scale_x_datetime(limits = c(as.POSIXct("2041-07-16 23:15:00
#T"), as.POSIXct("2041-07-17 02:15:00")))
 
ggplotly(trans_gb_s_timeseries)

##trans_db_n plot
trans_db_n_timeseries <- trans_db_n %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  facet_wrap(~aspect) + 
  labs(x = "Datetime", y = "Temp (C)", title = "trans_db_n")
  #scale_x_datetime(limits = c(as.POSIXct("2041-07-16 23:15:00
#T"), as.POSIXct("2041-07-17 02:15:00")))
 
ggplotly(trans_db_n_timeseries)

##trans_db_s plot
trans_db_s_timeseries <- trans_db_s %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  facet_wrap(~aspect) + 
  labs(x = "Datetime", y = "Temp (C)", title = "trans_db_s")
  #scale_x_datetime(limits = c(as.POSIXct("2041-07-16 23:15:00
#T"), as.POSIXct("2041-07-17 02:15:00")))
 
ggplotly(trans_db_s_timeseries)

#FOLLOWING GRAPHS ARE NOT AS HELPFUL
#graph
ggplot(trans_b_aspects, aes(x = aspect, fill = aspect)) +
  geom_bar() +
  labs(title = "Aspect vs Temperature",
       x = "Aspect",
       y = "Count") +
  theme_minimal() +
  facet_wrap("sensor_id")

#timeseries plots by tree
trans_timeseries <- trans_b_aspects %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  facet_wrap(~sensor_name) + 
  labs(x = "Datetime", y = "Temp (C)")

print(trans_timeseries)

#timeseries plot of just aspect... not as helpful
trans_b_aspect_timeseries <- trans_b_aspects %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp)) + facet_wrap("aspect")

plot(trans_b_aspect_timeseries)

#histogram
trans_b_histogram <- trans_b_aspects %>%
  ggplot(aes(x = temp)) +
  geom_histogram() + facet_wrap( ~ aspect, scale = "free_x") +
  labs(title = "Histogram of Temperature Across Sites",
       x = "Temperature (C)",
       y = "Frequency") +
  theme_minimal()
    # geom_vline(aes(xintercept = mean_temp), price_stats, color = "red", linewidth = 2)

print(trans_b_histogram)

trans_gb_n_timeseries <- trans_gb_n %>%
  ggplot() +
  geom_line(aes(x = datetime, y = temp, color = aspect)) +
  facet_wrap(~sensor_name) + 
  labs(x = "Datetime", y = "Temp (C)")
  
print(trans_gb_n_timeseries)
```

### 105E Trans_b Data Exploration

```{r}
#trans_north_GB tree temp data
#view the dataset - must use data() function to bring it into the environment
trans_n_gb_temp <- trans_b_aspects %>%
  filter(aspect == "north") %>%
  filter(tree_name %in% c("gb_n", "gb_s")) %>%
  drop_na()
  #filter(aspect == "north" & datetime >= "2023-12-07 00:00:00" & datetime <= "2024-01-10 00:00:00")

#glimpse is a different way to view dataset, a bit more clear
glimpse(trans_n_gb_temp)

#create a new variable to test 
trans_n_gb_clean<- trans_n_gb_temp %>%
  filter(aspect == "north")

#visualize the relationship
trans_n_gb_clean %>%
  ggplot(aes(x = tree_name, y = temp)) +
  geom_boxplot()

#check p-value and assumptions
trans_n_gb_clean %>%
  levene_test(temp ~ tree_name)
#this gives me a p-value of 0.102, which is not significant, which means there is not a difference in variances noted... 
```

```{r}
#pers_n_db tree temp comparison
trans_n_db_temp <- trans_b_aspects %>%
  filter(aspect == "north",tree_name %in% c("db_n", "db_s")) %>%
  drop_na()
  #filter(aspect == "north" & datetime >= "2023-12-07 00:00:00" & datetime <= "2024-01-10 00:00:00")

glimpse(trans_n_db_temp)

trans_n_db_clean<- trans_n_db_temp

trans_n_db_clean %>%
  ggplot(aes(x = tree_name, y = temp)) +
  geom_boxplot()

#check p-value and assumptions
trans_n_db_clean %>%
  levene_test(temp ~ tree_name)

#not a significant p-value, 0.800
```

```{r}
#trans_s_gb tree temp data
trans_s_gb_temp <- trans_b_aspects %>%
  filter(aspect == "south") %>%
  filter(tree_name %in% c("gb_n", "gb_s")) %>%
  drop_na()
  #filter(aspect == "south" & datetime >= "2023-12-07 00:00:00" & datetime <= "2024-01-10 00:00:00")

#glimpse is a different way to view dataset, a bit more clear
glimpse(trans_s_gb_temp)

#create a new variable to test 
trans_s_gb_clean<- trans_s_gb_temp

#visualize the relationship
trans_s_gb_clean %>%
  ggplot(aes(x = tree_name, y = temp)) +
  geom_boxplot()

#check p-value and assumptions
trans_s_gb_clean %>%
  levene_test(temp ~ tree_name)
#gives a p-value of 0.601

hist(trans_s_gb_clean$temp)

#this part of test did not work...
trans_s_gb_clean %>%
  levene_test(log(temp) ~ tree_name)

#performing t-test on non-transformed variable
welch_t_test_result <- trans_s_gb_clean %>%
  t_test(temp ~ tree_name, var.equal = FALSE)

# Print the result
print(welch_t_test_result)
#p-value is not significant - 0.758 value
```

```{r}
#trans_s_db tree temp comparison
trans_s_db_temp <- trans_b_aspects %>%
  filter(aspect == "south" & tree_name %in% c("db_n", "db_s")) %>%
  drop_na()
  #filter(aspect == "south" & datetime >= "2023-12-07 00:00:00" & datetime <= "2024-01-10 00:00:00")

glimpse(trans_s_db_temp)

trans_s_db_clean<- trans_s_db_temp

trans_s_db_clean %>%
  ggplot(aes(x = tree_name, y = temp)) +
  geom_boxplot()

print(trans_s_db_clean)

#check p-value and assumptions
trans_s_db_clean %>%
  levene_test(temp ~ tree_name)
#p-value is not significant - 0.273

trans_s_db_clean %>%
  levene_test(log(temp) ~ tree_name)
#p-value is not significant, 0.642
```

Plotting GB_N vs GB_S for transitional data

```{r}
#creating a new dataframe with name changes to sensors
trans_n_gb_new <- trans_n_gb_temp %>%
  mutate(tree_name = case_when(
    tree_name == "gb_n" ~ "n_gb_n",
    tree_name == "gb_s" ~ "n_gb_s",
    TRUE ~ tree_name  # Keep the original value if it doesn't match the conditions
  ))

trans_s_gb_new <- trans_s_gb_temp %>%
  mutate(tree_name = case_when(
    tree_name == "gb_n" ~ "s_gb_n",
    tree_name == "gb_s" ~ "s_gb_s",
    TRUE ~ tree_name  # Keep the original value if it doesn't match the conditions
  ))

#facet wrapped plots by tree
trans_n_s_gb_plot <- ggplot() +
  geom_line(data = trans_n_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "North"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = trans_s_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "South"), alpha = 0.3) +
  scale_color_manual(values = c("blue", "red", "blue", "red"), name = NULL) +
  scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans GB North and South Sensors")
#facet_wrap(~tree_name)

ggplotly(trans_n_s_gb_plot)
```

Plotting DB_N vs DB_S for transitional sites

```{r}
#creating a new dataframe with name changes to sensors
trans_n_db_new <- trans_n_db_temp %>%
  mutate(tree_name = case_when(
    tree_name == "db_n" ~ "n_db_n",
    tree_name == "db_s" ~ "n_db_s",
    TRUE ~ tree_name  # Keep the original value if it doesn't match the conditions
  ))

trans_s_db_new <- trans_s_db_temp %>%
  mutate(tree_name = case_when(
    tree_name == "db_n" ~ "s_db_n",
    tree_name == "db_s" ~ "s_db_s",
    TRUE ~ tree_name  # Keep the original value if it doesn't match the conditions
  ))

#plotting
trans_n_s_db_plot <- ggplot() +
  geom_line(data = trans_n_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "North"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = trans_s_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "South"), alpha = 0.3) +
scale_color_manual(values = c("blue", "red", "blue", "red"), name = NULL) +
  scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans DB North and South Sensors")
#facet_wrap(~tree_name)

ggplotly(trans_n_s_db_plot)
```

Plotting Trans N GB vs DB

```{r}
trans_n_gb_vs_db_plot <- ggplot() +
  geom_line(data = trans_n_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "North"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = trans_n_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "South"), alpha = 0.3) +
  scale_color_manual(values = c("blue", "red", "blue", "red"), name = NULL) +
  #scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans North GB vs DB")

ggplotly(trans_n_gb_vs_db_plot)
```

Plotting Trans S GB vs DB

```{r}
trans_s_gb_vs_db_plot <- ggplot() +
  geom_line(data = trans_s_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "North"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = trans_s_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "South"), alpha = 0.3) +
  scale_color_manual(values = c("blue", "red", "blue", "red"), name = NULL) +
  #scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans South GB vs DB")

ggplotly(trans_s_gb_vs_db_plot)
```

### Comparing Trans vs Pers Sensors

```{r}
#do stacked plots like I did for radiation in pers_burned area
```

Dead burn comparison

```{r}
#plotting trans vs persistent db sesnors
trans_vs_pers_n_db <- ggplot() +
  geom_line(data = trans_n_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "Transitional"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = pers_n_db_new, aes(x = datetime, y = temp, color = tree_name, linetype = "Persistent"), alpha = 0.3) +
scale_color_manual(values = c("blue", "red", "blue", "red"), name = NULL) +
  scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans vs Pers DB North Sensors")
#facet_wrap(~tree_name)

ggplotly(trans_vs_pers_n_db)
```

Green burn comparison

```{r}
#plotting trans vs persistent gb sesnors
trans_vs_pers_n_gb <- ggplot() +
  geom_line(data = trans_n_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "Transitional"), linetype = "dashed", alpha = 0.5) +
  geom_line(data = pers_n_gb_new, aes(x = datetime, y = temp, color = tree_name, linetype = "Persistent"), alpha = 0.3) +
scale_color_manual(values = c("blue", "red", "blue", "red"), name = NULL) +
  scale_linetype_manual(values = c("North" = "dashed", "South" = "solid"), name = "Region") +
  labs(x = "Datetime", y = "Temp (C)", title = "Trans vs Pers GB North Sensors")
#facet_wrap(~tree_name)

ggplotly(trans_vs_pers_n_gb)
```
